Correlations between gender and research topics at three major archaeology conferences

### 2021-03-15 analysis/paper/paper.Rmd
added basic plot of data

library(gender)
lynne_gender <-  gender("Lynne")

@@

SAA18_gm <- readr::read_rds(here::here("analysis/data/derived-data/SAA18_gender_model.rds")) 
EAA18_gm <- readr::read_rds(here::here("analysis/data/derived-data/EAA18_gender_model.rds"))
CAA18_gm <- readr::read_rds(here::here("analysis/data/derived-data/CAA18_gender_model.rds"))

@@

count_gendered_CAA18 <- nrow(CAA18_gender)
perc_gendered_CAA18 <- round(nrow(CAA18_gender) / CAA18_row, 2) * 100
female_count_CAA18 <- 
  length(which(CAA18_gender$gender == "female"))
male_count_CAA18 <- 
  length(which(CAA18_gender$gender == "male"))
female_to_male_CAA18 <- 
  round(female_count_CAA18 / male_count_CAA18, 1)
missing_gender_CAA18 <-
  CAA18_row - (female_count_CAA18 + male_count_CAA18)
```
Overall we see similar women-to-men ratios for the SAA (`r female_to_male_SAA18`) and EAA (`r female_to_male_EAA18`), but much lower for the CAA (`r female_to_male_CAA18`). Even before we look into the gender preferences for specific topics within each conference, we already have a hint that computational topics may be less preferred by women than by men.

@@

>>>>

library(gender)
lynne_gender <-  gender("Lynne")
```

@@

SAA18_gm <- 
  readr::read_rds(here::here("analysis/data/derived-data/SAA18_gender_model.rds"))

EAA18_gm <- 
  readr::read_rds(here::here("analysis/data/derived-data/EAA18_gender_model.rds"))

CAA18_gm <- 
  readr::read_rds(here::here("analysis/data/derived-data/CAA18_gender_model.rds"))

@@

count_gendered_CAA18 <- nrow(CAA18_gender)

perc_gendered_CAA18 <- round(nrow(CAA18_gender) / CAA18_row, 2) * 100

female_count_CAA18 <- 
  length(which(CAA18_gender$gender == "female"))

male_count_CAA18 <- 
  length(which(CAA18_gender$gender == "male"))

female_to_male_CAA18 <- 
  round(female_count_CAA18 / male_count_CAA18, 1)

missing_gender_CAA18 <-
  CAA18_row - (female_count_CAA18 + male_count_CAA18)
```

```{r}
# some basic visualisations to show the data
list_confs <- 
  list(SAA18_gender %>% 
              mutate(session = as.character(session)),
       CAA18_gender,
       EAA18_gender %>% 
              mutate(session = as.character(session))
       )

names(list_confs) <- c("SAA", "CAA", "EAA")

all_conf_tbl <- 
  bind_rows(list_confs,
            .id = "conf")

# how many papers per conf?
ggplot(all_conf_tbl) +
  aes(conf,
      fill = gender) +
  geom_bar(position = "dodge") +
  theme_minimal(base_size = 16) +
  theme(legend.position = c(.2,.60)) +
  labs(x = "", 
       y = "number of papers presented in 2018")
```

Overall we see similar women-to-men ratios for the SAA (`r female_to_male_SAA18`) and EAA (`r female_to_male_EAA18`), but much lower for the CAA (`r female_to_male_CAA18`). Even before we look into the gender preferences for specific topics within each conference, we already have a hint that computational topics may be less preferred by women than by men.

@@

A key limitation of this study is our concept of gender, and how we measure it. Like many social science studies, we have measured gender assuming it has two binary categories. However, we recognise that many cultures have long included more than two genders [@wilson1996we; @graham2004s], and there is a growing awareness of this in Western cultures also. A better measure of gender would represent it as a multifaceted spectrum [@tobin2010intrapsychics]. The problem with our binary measure is that it does not reflect our current understanding of gender, forces people in misclassified categories, and is hostile to public acceptance and advocacy for transgender and nonbinary individuals. Our instrument for collecting gender data only returns two categories, so this has limited potential for overcoming this limitation. Instead, we recommend that conference organisers collect gender information directly from presenters using inclusive gender measures. For example, by providing nonbinary options for presenters on registration forms, and by asking about gender as an open‐ended question for participants to self-identify. We recommend that future work on gender in archaeology avoid 'othering' language when describing results. For example, reporting a sample as '150 participants (48% women; 49% men; 3% other)' violates ethical standards because such wording implies that binary gender is normal or appropriate, whereas trans and nonbinary gender is not (it is 'other'). Our hope is that our study might be the last one to  discuss gender as a binary.
